---
title: "Explorative Analysis - item wrenexp"
subtitle: "“How worried are you that energy may be too expensive for many people in [country]?"
author: "Julius Fenn"
date: "3.3.2024"
output:
  html_document:
    toc: yes
    toc_depth: 3
    number_sections: yes
  pdf_document:
    toc: yes
    toc_depth: '3'
  word_document:
    toc: yes
    toc_depth: '3'
---


```{css, echo=FALSE}
.CSSoutput {
  background-color: white;
  border: 1px solid black;
}
```
```{r setup chunks, include=FALSE}
knitr::opts_chunk$set(class.source = "CSSoutput",
                      eval = TRUE, echo = TRUE, include = TRUE,      
                      fig.align='center', fig.show='asis',
                      size='footnotesize')

```


```{r setup, include=FALSE}
# reset R environment
rm(list=ls(all=TRUE))
graphics.off()

################
# install and load packages
################
#  if packages are not already installed, the function will install and activate them
usePackage <- function(p) {
  if (!is.element(p, installed.packages()[,1]))
    install.packages(p, dep = TRUE, repos = "http://cran.us.r-project.org")
  require(p, character.only = TRUE)
}




usePackage("haven") # load SPSS, ... data 
usePackage("tidyverse") # data cleaning and summarizing
usePackage("psych") # psychometric analysis (EFA)

## psychometric analysis
usePackage("moments") # skewness, kurtosis

## outputs
usePackage("stargazer") # create tables

usePackage("ggstatsplot")

usePackage("MASS")



rm(usePackage)
```

```{r loaddata, include=FALSE}
## change working directory
setwd("data")
## load data
dat <- haven::read_spss(file = "ESS-Data-Wizard-subset-2023-04-26.sav")


## data preparation


# reverse coded items



## create mean scores

setwd("..")
```





# Results

## Descriptive statistics

*overview data:*
```{r}
head(dat)
dim(dat)
```

*variable of interest (wrenexp):*

distribution variable:
```{r}
attr(x = dat$wrenexp, which = "label")
attr(x = dat$wrenexp, which = "labels")



barplot(table(dat$wrenexp))
round(x = table(dat$wrenexp) / nrow(dat) * 100, digits = 2)
```

mean differences over countries: 


```{r}
tmp <- dat %>%
  group_by(cntry) %>%
  summarise(N = n(), mean = mean(x = wrenexp, na.rm = TRUE))
tmp[order(tmp$mean),]


boxplot(dat$wrenexp ~ dat$cntry)


tmp <- dat
tmp$wrenexp <- as.numeric(tmp$wrenexp)
tmp <- tmp[tmp$cntry %in% c("AT", "CH", "CZ", "DE", " ES", "FR"),] # only for subset of countries

p <- ggbetweenstats(data = tmp,
  x    = cntry,
  y    = wrenexp,
  type = "p" # parametric
)
p
```
### swiss specific

```{r}
dat_CH <- dat
dat_CH <- dat_CH[dat_CH$cntry == "CH",]

barplot(table(dat_CH$wrenexp))
round(x = table(dat_CH$wrenexp) / nrow(dat_CH) * 100, digits = 2)
summary(dat_CH$wrenexp)


sum(dat_CH$wrenexp >= 4, na.rm = TRUE) / nrow(dat_CH)
```

> only about 12% are Very worried  or extremely worried 

## (stepwise) logistic regression

see: https://www.methodenberatung.uzh.ch/de/datenanalyse_spss/zusammenhaenge/lreg.html
stepwise procedure see: http://www.sthda.com/english/articles/36-classification-methods-essentials/150-stepwise-logistic-regression-essentials-in-r/

```{r}
dat_logistic <- dat
dat_logistic <- dat_logistic[dat_logistic$cntry == "CH",]
dat_logistic <- dat_logistic[, colSums(x = is.na(dat_logistic)) != nrow(dat_logistic)]
dat_logistic <- dat_logistic[, colSums(x = is.na(dat_logistic)) < 1000]
dat_logistic <- dat_logistic[!is.na(dat_logistic$wrenexp), ]

dat_logistic$wrenexp_dummy <- ifelse(test = dat_logistic$wrenexp >= 4, yes = 1, no = 0) 
table(dat_logistic$wrenexp_dummy)

dat_logistic <- dat_logistic %>%
  dplyr::select(netusoft:impfun, wrenexp_dummy, wrenexp)

dat_logistic <- dat_logistic %>%
  dplyr::select(!c(netustm, prtvtfch, badge))



var_IVs <- colnames(dat_logistic)[1:20]
dat_logistic_out <- dat_logistic[, c("wrenexp_dummy", var_IVs, "wrenexp")]
dat_logistic_out <- na.omit(dat_logistic_out)

m <- glm(wrenexp_dummy ~., data = dat_logistic_out[, 1:20], family = binomial)
summary(m)

step.model <- m %>% stepAIC(trace = FALSE)
summary(step.model)
coef(step.model) 
# to interpret the params: 
 exp(coef(step.model)) - 1
 
 
# trstlgl
# Trust in the legal system
# Value 	Category
# 0 	No trust at all
# 1 	1
# 2 	2
# 3 	3
# 4 	4
# 5 	5
# 6 	6
# 7 	7
# 8 	8
# 9 	9
# 10 	Complete trust
 dat_logistic_out %>%
   group_by(trstlgl) %>%
   summarise(N = n(), mean = mean(x = wrenexp), sd = sd(x = wrenexp))
 
# if "Trust in the legal system" increases by one unit the probability to be "Very worried" or "Extremely worried" about energy prices decreases by around 10%
 

# psppsgva 
# Political system allows people to have a say in what government does
# Value 	Category
# 1 	Not at all
# 2 	Very little
# 3 	Some
# 4 	A lot
# 5 	A great deal
  dat_logistic_out %>%
   group_by(psppsgva) %>%
   summarise(N = n(), mean = mean(x = wrenexp), sd = sd(x = wrenexp))
  
# if "Political system allows people to have a say in what government" increases by one unit the probability to be "Very worried" or "Extremely worried" about energy prices decreases by 28%
```


## (stepwise) ordinal logistic regression

ordinal logistic regression see: https://stats.oarc.ucla.edu/r/dae/ordinal-logistic-regression/
stepwise procedure see: http://www.sthda.com/english/articles/36-classification-methods-essentials/150-stepwise-logistic-regression-essentials-in-r/


```{r}
# dat_logistic <- dat
# dat_logistic <- dat_logistic[dat_logistic$cntry == "CH",]
# dat_logistic <- dat_logistic[, colSums(x = is.na(dat_logistic)) != nrow(dat_logistic)]
# dat_logistic <- dat_logistic[, colSums(x = is.na(dat_logistic)) < 1000]
# dat_logistic <- dat_logistic[!is.na(dat_logistic$wrenexp), ]
# 
# dat_logistic <- dat_logistic %>%
#   dplyr::select(netusoft:impfun)
# 
# dat_logistic <- dat_logistic %>%
#   dplyr::select(!c(netustm, prtvtfch, badge))
# 
# 
# dat_logistic$wrenexp <- factor(dat_logistic$wrenexp, order = TRUE)
# 
# 
# var_IVs <- colnames(dat_logistic)[1:20]
# dat_logistic_out <- dat_logistic[, c("wrenexp", var_IVs)]
# dat_logistic_out <- na.omit(dat_logistic_out)
# 
# m <- polr(formula = wrenexp ~ ., data = dat_logistic_out, Hess=TRUE)
# summary(m)
# 
# step.model <- m %>% stepAIC(trace = FALSE)
# coef(step.model)
# 
# 
# 
# ## get labels of all significant parameters
# for(n in names(coef(step.model))){
#   print(n)
#   print(attr(x = dat[[n]], which = "label"))
#     print(attr(x = dat[[n]], which = "labels"))
#   cat("\n")
# }
```


### swiss specific

mean differences between parties

```{r}
dat_CH <- dat
dat_CH <- dat_CH[dat_CH$cntry == "CH",]


tmp <- dat_CH %>%
  group_by(prtvtfch) %>%
  summarise(N = n(), mean = mean(x = wrenexp, na.rm = TRUE))
tmp[order(tmp$mean),]
```
