---
title: "What socio-demographic factors influence European's willingness to commit to individual climate change adaptation and mitigation measures"
author: "Jana Kalmbach"
date: "2022-03-28"
output:
  bookdown::html_document2:
    toc: yes
    toc_depth: 3
    number_sections: yes
  pdf_document:
    toc: yes
    toc_depth: '3'
  word_document:
    toc: yes
    toc_depth: '3'
bibliography: References.bib
biblio-style: apalike
link-citations: yes
---


```{css, echo=FALSE}
.CSSoutput {
  background-color: white;
  border: 1px solid black;
}
```
```{r setup chunks, include=FALSE}
knitr::opts_chunk$set(class.source = "CSSoutput",
                      eval = TRUE, echo = TRUE, include = TRUE,      
                      fig.align='center', fig.show='asis',
                      size='footnotesize')

```


```{r setup, include=FALSE}
# reset R environment
rm(list=ls(all=TRUE))
graphics.off()

################
# install and load packages
################
#  if packages are not already installed, the function will install and activate them
usePackage <- function(p) {
  if (!is.element(p, installed.packages()[,1]))
    install.packages(p, dep = TRUE, repos = "http://cran.us.r-project.org")
  require(p, character.only = TRUE)
}
usePackage("haven") # load SPSS, ... data 
usePackage("tidyverse") # data cleaning and summarizing
usePackage("psych") # psychometric analysis (EFA)
usePackage("dplyr") # to clean data
usePackage("readxl")
usePackage("ggplot2")
usePackage("ggpubr")
## psychometric analysis
usePackage("moments") # skewness, kurtosis
usePackage("stargazer") # create tables
rm(usePackage)
```

```{r loaddata, include=FALSE}
# change working directory
setwd('data')

## load data
dat <- read_excel("eib-climate-survey-2021-2022-bdd-all-countries-results-pr1-2.xlsx")

########################
### Data preparation  ##
########################

# select only a couple of countries of interest
dat <- dplyr::filter(dat, grepl('Germany|Sweden|France|Finland|Portugal', country))

# rename column headers
colnames(dat)[colnames(dat) %in% c("sd2", "sd8", "sd1", "sd3", "sd6b", "sd7", "Q25")] <- c("age", "children 18 yo", "gender", "detailed occupation", "left-right", "conservative", "pop.density")

#remove unnecessary columns
dat$record <- NULL # participant number
dat$uuid <- NULL # participant id
dat$language <- NULL # language of questionnaire
dat$mpq <- NULL
dat$mpq2 <- NULL
dat$sd6 <- NULL
dat <- subset (dat, select = -c(qc1:agerecode))
dat <- subset (dat, select = -c(sd4at:sd5cn))
dat$age <- NULL
dat$detailed.occupation <- NULL
dat$conservative <- NULL
dat <- subset (dat, select = -c(Q1r1:Q1r13))
dat$Q2 <- NULL
dat <- subset (dat, select = -c(Q4:Q5))
dat$Q10 <- NULL
dat$wght1 <- NULL
dat$wght2 <- NULL
dat <- dat[-c(1), ] # delete row with survey questions from dataset. For more details on the questions, see codebook in the data folder.

# rename some answer options
dat$occupation <- recode(dat$occupation, "Currently not economically active (retired, students, homemakers and others)" = 'Not active', "Net economically active"  = 'Active', "Unemployed" = 'Unemployed')
dat$pop.density <- recode(dat$pop.density, "A city/metropolis" = 'City', "A rural area"  = 'Rural', "A small or medium-sized town" = 'Medium')

# change character values to categorical
dat <- as.data.frame(unclass(dat),         # Convert all columns to factor
                       stringsAsFactors = TRUE)

############################
##  Prepare climate score ##
############################

# store the prepared data for later analysis 
factor_dat <- dat 

# change data to numeric
dat[sapply(dat, is.factor)] <- data.matrix(dat[sapply(dat, is.factor)])

# clean data:

# political affiliation: do not wish to reply -> turn to Nas, ensure that order is correct
dat$left.right[dat$left.right == 11] <- NA
dat$left.right[dat$left.right == 2] <- 12
dat$left.right <- dat$left.right - 1
dat$left.right[dat$left.right == 11] <- 10
dat$left.right[dat$left.right == 0] <- 1

# exclude questions that are not relevant for the personal climate score
dat$Q6 <- NULL
dat$Q7 <- NULL
dat <- subset (dat, select = -c(Q9r1:Q9r4))
# dat$Q13 <- NULL
dat <- subset (dat, select = -c(Q16r1:Q16r2))
dat$Q19 <- NULL

# in Q11, the scale is wrong
# this is certainly not the most elegant way to reorder the score rating, but it works
dat$Q11[dat$Q11 == 4] <- 5
dat$Q11[dat$Q11 == 3] <- 4
dat$Q11[dat$Q11 == 5] <- 3

# in Q12, the scale is wrong
dat$Q12[dat$Q12 == 4] <- 5
dat$Q12[dat$Q12 == 3] <- 4
dat$Q12[dat$Q12 == 5] <- 3

# in Q 14 the scale is wrong
dat$Q14r1[dat$Q14r1 == 4] <- 5
dat$Q14r1[dat$Q14r1 == 3] <- 4
dat$Q14r1[dat$Q14r1 == 5] <- 3

dat$Q14r2[dat$Q14r2 == 4] <- 5
dat$Q14r2[dat$Q14r2 == 3] <- 4
dat$Q14r2[dat$Q14r2 == 5] <- 3

dat$Q14r3[dat$Q14r3 == 4] <- 5
dat$Q14r3[dat$Q14r3 == 3] <- 4
dat$Q14r3[dat$Q14r3 == 5] <- 3

dat$Q14r4[dat$Q14r4 == 4] <- 5
dat$Q14r4[dat$Q14r4 == 3] <- 4
dat$Q14r4[dat$Q14r4 == 5] <- 3

dat$Q14r5[dat$Q14r5 == 4] <- 5
dat$Q14r5[dat$Q14r5 == 3] <- 4
dat$Q14r5[dat$Q14r5 == 5] <- 3

# reverse Q18
dat$Q18 <- 5 - dat$Q18
```


# Introduction
Awareness and concern about climate change did globally rise in the past decades [@driscoll_assessing_2019]. However, while some young adolescents go on the streets on Fridays to demonstrate for a need for immediate action, others deny that climate change even exists. Sociological research has examined what socio-demographic factors are potentially related to certain attitudes towards climate change. Researchers suggest correlations with age [@krkoska_lorencova_perception_2019; @poortinga_climate_2019], gender [@korkmaz_public_2018], race [@ballew_does_2020] income and education [@ballew_does_2020; @poortinga_climate_2019], unemployment [@benegal_impact_2018], and having adolescent children or grandchildren [@lawson_children_2019]. Additionally, research stresses a high correlation between climate change attitudes and political orientation. For instance, @huber_role_2020 found a significant correlation between climate skepticism and populist worldviews in the US. Additionally, @krange_cool_2019 state that conservative, white males tend to hold the highest share of climate skeptics in Norway.  In his metanalysis, @hornsey_meta-analyses_2016 concluded that attitudes towards climate change are better explained by political orientation, ideologies, and worldviews than other socio-demographic factors. However, @lewis_cross-national_2019 criticize than many researchers focus solely on the US, while party identification and political ideology seem to matter much less across the globe. 


To assess the situation in Europe, this research aims to explore **which socio-demographic factors influence European’s willingness to personally commit to climate change adaptation and mitigation measures**. Taken measures include for instance the restriction of individual consumption or the conscious decision to consume more regional and environmentally friendly products. Based on data of the climate survey of the European Investment bank conducted in 2021, climate scores are calculated of all German, French, Swedish, Portuguese, and Finish participants. A multiple linear regression is applied to test for correlations between climate score and socio-demographic factors. Additionally, socio-demographic factors of 5% of the participants achieving the highest and lowest climate score are compared. Even though this study has some statistical and theoretical limitations, it nevertheless contributes to the research about potential predictors for certain attitudes towards climate change. Such research might be relevant to better target climate change communication or policies. 

# Methods
For this study, data is retrieved from the European Investment Bank, that has conducted a climate survey every year since 2018 [@european_investment_bank_climate_nodate]. Since the different rounds are not directly comparable due to varying questions, this study focuses only on the last round conducted in August and September 2021. Interviewees participated at the self-administered survey online in all 27 EU countries, the UK, China, and the US. Respondents were randomly selected among nationally representative panels. The total sample includes 30.662 participants aged 15 and older. Depending on the amount of inhabitants per country, the sample size ranged from 500 to 2.000, except Malta (n=200) and Cyprus (n=462). The representativity of the data was ensured by the quota method and samples in each country were weighted according to gender, age, occupation, and region. To reduce the amount of data, this study focuses only on Germany, France, Sweden, Portugal, and Finland (n = 6.999). The countries are chosen for their major contribution to European emissions (Germany, France) [@european_parliament_greenhouse_2018], because they were specifically mentioned in the literature [@brink_citizen_2019], or because of their location in Europe. 

## Hypothesis

*H0: Some socio-demographic factors could potentially serve as predictors for European's willingness to commit to climate change adaptation and mitigation measures.*

*H1: Socio-demographic factors cannot serve as predictors for European's willingness to commit to climate change adaptation and mitigation measures.*

To test the null-hypothesis, survey questions that point towards individuals’ willingness for climate change adaptation and mitigation measures are determined. Depending on the answer of participants to these questions, their individual climate score is calculated. Then, a multiple linear regression is conducted to explore whether a high score correlates with country, age, gender, occupation, political affiliation, or population density of home.
For further analysis, the number of participants that deny anthropogenic climate change are identified and their mean climate score is compared to the entire sample. Moreover, the socio-demographic factors of the 175 individuals (5% of the sample) with the highest and lowest climate scores respectively are compared.

# Results

First, the climate score is calculated.

```{r}
climate_score <- dat %>%
  select(c(Q3:Q18)) %>%
  rowSums()


dat %>%
  select(c(Q11:Q13))
```

```{r include = FALSE}
# add climate score column to initial data set
dat$climate_score <- climate_score
factor_dat$climate_score <- climate_score
```

## Descriptive analysis of climate score
Based on the following descriptive analysis (see Table \@ref(tab:describc)), we can assume that the climate score is relatively normally distributed (skewness = -0.34, kurtosis = -0.46). The mean score for all participants (n = 6999) is 41.35 points with a median of 42 points, and a standard deviation of 7.62. The highest score achieved were 59 points (n = 1) and the lowest were 19 points (n = 2).   

```{r describc}
tmp <- describe(climate_score)  # create a descriptive summary
tmp$var <- var(climate_score)   # add the variance
#rename row and column names
rownames(tmp) <- "Climate score"
colnames(tmp) <- c("vars",	"n",	"mean",	"standard deviation",	"median",	"trimmed mean",	"mad",	"min",	"max", "range",	"skewness",	"kurtosis",	"se",	"variance")
# display the descriptive summary as table
knitr::kable(tmp, caption = "Summary of climate score", digits = 3)
```

In Figure \@ref(fig:fig1), the climate score is represented as a histogram and a boxplot. 

```{r fig1, fig.cap = "Histogram and boxplot representation of climate score of the sample"}
par(mfrow = c(1,2)) # 2 plots in 1 column
hist(climate_score,  main="Histogram of climate score",
     xlab="Achieved climate score",
     ylab="Frequency")
boxplot(climate_score, main="Boxplot of climate score",
        ylab="Climate score")
```

As can be seen from Table \@ref(tab:describc) and Figure \@ref(fig:fig1), there are more individuals that have higher climate scores than those with lower ones. We can express this by categorizing the participants willingness to commit to climate change adaptation and mitigation measures into below average, medium, and above average (see Table \@ref(tab:cat)).  

```{r cat}
cat <- table(cut(x = climate_score, breaks = 3,
                             labels = c("below average", "medium", "above average")))
# add proportinal share to table
prop <- cat %>% prop.table %>% round(3)
climate_score_cat <- as.data.frame(cat)
climate_score_cat$prop <- prop
# rename columns
colnames(climate_score_cat) <- c("Climate score performance", "Number of individuals", "Proportions")
# display as table
knitr::kable(climate_score_cat, caption = "Climate score distribution among individuals", digits = 3)
```


## Linear regression exploring socio-demographic factors

A multiple linear regression is performed to test whether the climate score correlates with age, gender, country, children under 18 years-old, occupation, political affiliation, and population density of home.

```{r, results='asis'}
lm1 <- lm(formula = climate_score ~ origage + gender + country + children.18.yo + occupation + left.right + pop.density, data = dat)
stargazer(lm1, type = "html") # display as table
```

The multiple linear regression shows a significant correlation between climate score and gender, children under 18 years old, political affiliation, and population density. The other factors are not significantly correlated. The R squared value is 0.029. It should be mentioned that 707 values were excluded because the data about political affiliation contained unknown values.

In a next step, the correlation relationships are displayed graphically. This could be done by making boxplots comparing each two variables respectively. Figure \@ref(fig:boxplot) shows the climate score by gender, having children or grandchildren, political affiliation, and population density. The mean is displayed as red line. The plots show that participants that are female, have children or grandchildren younger than 18 years-old, who are left-wing, and who live in a city or metropolis achieve higher climate scores in comparison. Additionally, it seems that there is a higher variation among people that are politically oriented towards the right with some achieving high climate scores and some really low ones. 

```{r boxplot, fig.cap = "Boxplots showing climate score by gender, having children or grandchildren under 18 years-old, political affilation, and population density of neighbourhood."}
par(mfrow = c(2,2)) # 2 plots per column, 2 columns
boxplot(climate_score ~ factor_dat$gender, ylab = "Climate score", xlab = "Gender", main = "Gender")
abline(h = 41.35, col = "red")  # add line of mean climate score of sample
boxplot(climate_score ~ factor_dat$children.18.yo, ylab = "Climate score", xlab = "Children or grandchildren younger than 18", main = "(Grand)children")
abline(h = 41.35, col = "red")
boxplot(climate_score ~ dat$left.right, ylab = "Climate score", xlab = "Political affiliation (1 = left, 10 = right)", main = "Political affiliation")
abline(h = 41.35, col = "red")
boxplot(climate_score ~ factor_dat$pop.density, ylab = "Climate score", xlab = "Population density", main = "Population density")
abline(h = 41.35, col = "red")

```

## Exploration of climate score, political affiliation, and other factors

Since political orientation is discussed in the literature as a new but essential predictor, further analysis will specifically focus on the relationships between climate score, political affiliation, and all other significant predictors.
The code for the plots displayed in Figure \@ref(fig:ggarange) is exemplarily displayed here for age. 
   
```{r}
# Age groups are added as categorical data
dat$lorigage <- factor_dat$origage

# relation between climate score, political affiliation, and age groups is plotted
orage <- ggplot(dat,
       aes(x = climate_score,
           y = left.right,
           color = lorigage)) +
  geom_point(size = .9,
             alpha = .3, color = "grey") +
  geom_smooth(method = "lm") +
  theme_bw() +
  labs(x = "Climate score",
       y = "Political affiliation",
       color = "Age groups")
```

```{r, include = FALSE}

#  relation between climate score, political affiliation, and gender
dat$lgender <- factor_dat$gender
gndr <- ggplot(dat,
       aes(x = climate_score,
           y = left.right,
           color = lgender)) +
  geom_point(size = .9,
             alpha = .3, color = "grey") +
  geom_smooth(method = "lm") +
  theme_bw() +
  labs(x = "Climate score",
       y = "Political affiliation",
       color = "Gender")

# relation between climate score, political affiliation, and children or grandchildren younger than 18
dat$lchildren <- factor_dat$children
chdrn <- ggplot(dat,
       aes(x = climate_score,
           y = left.right,
           color = lchildren)) +
  geom_point(size = .9,
             alpha = .3, color = "grey") +
  geom_smooth(method = "lm") +
  theme_bw() +
  labs(x = "Climate score",
       y = "Political affiliation",
       color = "(Grand)children")

# relation between climate score, political affiliation, and population density of neighbourhood
dat$lcity <- factor_dat$pop.density
cty <- ggplot(dat,
                aes(x = climate_score,
                    y = left.right,
                    color = lcity)) +
  geom_point(size = .9,
             alpha = .3, color = "grey") +
  geom_smooth(method = "lm") +
  theme_bw() +
  labs(x = "Climate score",
       y = "Political affiliation",
       color = "Living where?")
```

Repeating the same step for the other factors results in Figure \@ref(fig:ggarange). 

``` {r, ggarange, fig.cap= "A closer examination of the relationship between climate score, political affiliation, and other factors", warning = FALSE, message = FALSE}
# Plot all four graphs
figure <- ggarrange(orage, gndr, chdrn, cty,
                    ncol = 2, nrow = 2)
figure

```

Figure \@ref(fig:ggarange) shows tendencies between political affiliation, climate score and other factors. For instance, young adults (20-29 years) seem to achieve higher climate scores if they are slightly oriented towards the right end of the conservative spectrum and lower climate scores if they tend towards the left. Contrarily, for older participants (50+) the opposite seems to be the case. Concerning gender, both groups achieve higher scores if they tend towards the conservative to left-wing end. Additionally, females seem to have a higher tendency for the left spectrum than males. Interestingly, having young children or grandchildren seems to level out political affiliation, while participants without young family members achieve better climate scores if they are rather left-oriented. Lastly, people in rural areas and small or medium cities seem to be more environmentally conscious when being on the left end of the conservative spectrum. This relation is less strong but present for people living in big cities.   

## Comparative analysis of participants with low and high climate scores 

The 175 participants with the lowest and highest climate scores respectively are examined. The total of 250 participants equals to 5% of the entire dataset.

```{r}
# order participants based on climate scores
c_ordered <- factor_dat[order(factor_dat$climate_score),] 
# extract participants with lowest and highest scores
lowc <- c_ordered[1:175, ]
highc <- c_ordered[6825:6999,]
```

```{r include = FALSE}
# some data cleaning
highc <- subset (highc, select = -c(Q3:Q19))
lowc <- subset (lowc, select = -c(Q3:Q19))
highc$detailed.occupation <- NULL
lowc$detailed.occupation <- NULL
highc$cc <- as.factor("high c")
lowc$cc <- as.factor("low c")
```

```{r}
# combine into one data frame
cc <- rbind(lowc, highc)
# add id for each participant
cc$id <- 1:nrow(cc)
```

```{r include = FALSE}
# some more data cleaning (the left right scale is again wrong)
cc$left.right <- as.factor(cc$left.right)
cc$left.right <- ordered(cc$left.right, levels = c("1 - Very left-wing", "2", "3", "4", "5", "6", "7", "9", "10 - Very right-wing", "Do not wish to reply"))
```
A plot is computed for each socio-demographic variable. See plot 'a' as example:
```{r}
a <- ggplot(cc, aes(fill=country, x=cc, y=id)) + 
      geom_bar(position='fill', stat='identity')
```

```{r include = FALSE}
b <- ggplot(cc, aes(fill=gender, x=cc, y=id)) + 
  geom_bar(position='fill', stat='identity')
c <- ggplot(cc, aes(fill=occupation, x=cc, y=id)) + 
  geom_bar(position='fill', stat='identity')
d <- ggplot(cc, aes(fill=children.18.yo, x=cc, y=id)) + 
  geom_bar(position='fill', stat='identity')
e <- ggplot(cc, aes(fill=pop.density, x=cc, y=id)) + 
  geom_bar(position='fill', stat='identity')
f <- ggplot(cc, aes(fill=left.right, x=cc, y=id)) + 
  geom_bar(position='fill', stat='identity')
```

Figure \@ref(fig:plotall) and \@ref(fig:plotall2) graphically compare all socio-demographic factors for the 5% of the sample with the lowest and highest climate scores. It seems that Swedish and Finish participants generally tend to have lower climate scores. Contrarily, Portuguese and French people tend towards higher scores. For occupation, higher shares of net economically active and unemployed people achieve higher scores than lower ones. For all other factors, the plots mirror the previous findings. 

```{r, plotall, fig.cap = "Socio-demographic factors in comparison for participants with lowest and highest climate scores"}
figure1 <- ggarrange(a, b, c, d,
                    ncol = 2, nrow = 2)
figure1
```

```{r, plotall2, fig.cap = "Socio-demographic factors in comparison for participants with lowest and highest climate scores"}
figure2 <- ggarrange(e,f,
                    ncol = 2, nrow = 1)
figure2
```

## Additional analysis: Identification of climate deniers 

```{r include = FALSE}
# see if any participant did not answer question 6
table(is.na(factor_dat$Q6))
```

In Question 6 of the survey, participants are asked on their opinion why it is so difficult to solve the climate crisis. Two of the answering options let assume that participants do not believe in anthropogenic climate change. 

```{r}
cdeniers <- factor_dat[factor_dat$Q6 == 'There is no climate crisis, it is a hoax'|factor_dat$Q6 =='Because global warming is not due to human activities, so our actions won’t make a difference', ]
```

From Table \@ref(tab:deniers) we can see that 886 participants do not believe in anthropogenic climate change. This equals to more than 12% of the sample. The mean climate score is with 37.4 points almost 4 points lower than of the average participant. 

```{r deniers}
tmpp <- describe(cdeniers$climate_score) # create a summary 
knitr::kable(tmpp, caption = "Summary of climate score of climate deniers", digits = 3)
```

# Discussion
The results of this study mostly agree with the literature. The only difference is that there was no significant correlation between the climate score and age or occupation. Although the results should be seen with caution due to the simple analytical tools used, the null hypothesis can be accepted. Thus, there seems to be a significant correlation between some socio-demographic factors and the willingness to commit to climate change adaptation and mitigation measures in this sample. Nevertheless, there are some major limitations to this study that need to be addressed. 

## Limitations
First, only 0.03% of the variance can be explained by the linear regression model, which is comparably low for the social sciences. Moreover, ordinal scales are treated as numeric in this analysis. Instead, an ordinal logistic regression could have been performed, which is however difficult to interpret.  
Additionally, it needs to be questioned whether the climate score based on the survey questions realistically represents the willingness for climate change adaptation and mitigation measures for several reasons. First, the participants might have chosen answers based on how socially desirable they are or based on their self-estimation and not based on their real behavior. Additionally, some questions had to be removed from the survey because their answers were ambiguous to transfer into the score. Thirdly, the estimation which answer options are the most climate friendly might be wrong or there might be other confounding factors why participants chose this option over another. 
Moreover, not all socio-demographic factors were included in this study because they were not assessed in the survey. For instance, race, education, or income were not analyzed.  


# Conclusion
Overall, this study found a correlation between socio-demographic factors and Europeans’ willingness to commit to climate change adaptation and mitigation measures. Gender, political affiliation, having children or grandchildren younger than 18 years-old, and the population density of neighborhood were found to be predictors at a significant level. Generally, females with children or grandchildren that live in a city or metropolis and tend towards the left spectrum seem to be most willing to pursue a climate friendly lifestyle. Additionally, it seems that Portuguese and French are more willing to transform to a sustainable lifestyle than Fins and Swedes. In total, 12% of the sample does not believe in anthropogenic climate change. However, the results should be seen with caution because of statistical and theoretical limitations to this study. Generally, this study should also not be used to enhance existing stereotypes but might contribute to specifically address groups that do not yet thrive to reduce their emission footprint and to study their reasons and circumstances.  


# References
