---
title: "title"
author: "your name"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: yes
    toc_depth: 3
    number_sections: yes
  pdf_document:
    toc: yes
    toc_depth: '3'
  word_document:
    toc: yes
    toc_depth: '3'
bibliography: Library.bib
biblio-style: apalike
link-citations: yes
---


```{css, echo=FALSE}
.CSSoutput {
  background-color: white;
  border: 1px solid black;
}
```
```{r setup chunks, include=FALSE}
knitr::opts_chunk$set(class.source = "CSSoutput",
                      eval = TRUE, echo = TRUE, include = TRUE,      
                      fig.align='center', fig.show='asis',
                      size='footnotesize')

```


```{r packages, include=FALSE}
# reset R environment
rm(list=ls(all=TRUE)) # ! not needed in rmarkdown, but good practice
graphics.off() # ! not needed in rmarkdown

################
# install and load packages
################
#  if packages are not already installed, the function will install and activate them
usePackage <- function(p) {
  if (!is.element(p, installed.packages()[,1]))
    install.packages(p, dep = TRUE, repos = "http://cran.us.r-project.org")
  require(p, character.only = TRUE)
}

usePackage("haven") # load SPSS, ... data 
usePackage("tidyverse") # data cleaning and summarizing
usePackage("psych") # psychometric analysis (EFA)

## psychometric analysis
usePackage("moments") # skewness, kurtosis

## outputs
usePackage("stargazer") # create tables
rm(usePackage)
```

```{r additionalFunctions, include=FALSE}
################
# install and load packages
################
## change working directory
setwd("functions")
source(file = "getDescriptives.R", encoding = "UTF-8")
setwd("..") # ! not needed in rmarkdown, but good practice
```

```{r loaddata, include=FALSE}
################
# load and clean data
################
## change working directory
setwd("data")
## load data
dat <- haven::read_spss(file = "ESS1-9e01_1.sav")


## data preparation
dat$cntry <- NULL

# reverse coded items
dat$imsmetn <- 5 - dat$imsmetn 
dat$imdfetn <- 5 - dat$imdfetn
dat$impcntr <- 5 - dat$impcntr


## create mean scores
# > Immigration enriching economy, culture, place to live
sel_var <- c("imbgeco", "imueclt", "imwbcnt")
dat$mean_imEnrich <- rowMeans(x = dat[, sel_var], na.rm = TRUE)
# imbgeco	Immigration bad or good for country's economy	All rounds: Would you say it is generally bad or good for [country]'s economy that people come to live here from other countries?
# imueclt	Country's cultural life undermined or enriched by immigrants	All rounds: And, using this card, would you say that [country]'s cultural life is generally undermined or enriched by people coming to live here from other countries?
# imwbcnt	Immigrants make country worse or better place to live	All rounds: Is [country] made a worse or a better place to live by people coming to live here from other countries?
# >
sel_var <- c("imsmetn", "imdfetn", "impcntr")
dat$mean_imAllow <- rowMeans(x = dat[, sel_var], na.rm = TRUE)
# imsmetn	Allow many/few immigrants of same race/ethnic group as majority	ESS1, ESS2, ESS3, ESS4, ESS5, ESS6, ESS7, ESS8: Now, using this card, to what extent do you think [country] should allow people of the same race or ethnic group as most [country] people to come and live here?
# ESS9: Now, using this card, to what extent do you think [country] should allow people of the same race or ethnic group as most [country]'s people to come and live here?
# imdfetn	Allow many/few immigrants of different race/ethnic group from majority	All rounds: How about people of a different race or ethnic group from most [country] people?
# impcntr	Allow many/few immigrants from poorer countries outside Europe	All rounds: How about people from the poorer countries outside Europe?

# >
sel_var <- str_detect(string = colnames(dat), pattern = "^stf")
dat$mean_stf <- rowMeans(x = dat[, sel_var], na.rm = TRUE)
dat$sd_stf <- dat %>%
  select(matches(match = "^stf")) %>%
  apply(.,1, sd, na.rm = TRUE)
# stflife	How satisfied with life as a whole	All rounds: All things considered, how satisfied are you with your life as a whole nowadays? Please answer using this card, where 0 means extremely dissatisfied and 10 means extremely satisfied.
# stfeco	How satisfied with present state of economy in country	All rounds: On the whole how satisfied are you with the present state of the economy in [country]?
# stfgov	How satisfied with the national government	All rounds: Now thinking about the [country] government, how satisfied are you with the way it is doing its job?
# stfdem	How satisfied with the way democracy works in country	All rounds: And on the whole, how satisfied are you with the way democracy works in [country]?

setwd("..") # ! not needed in rmarkdown, but good practice
```





# Introduction

Migration 'mother of all political problems,' says German Interior Minister Horst Seehofer. Hort Seehofer's 2018 statement^[see <https://www.dw.com/en/migration-mother-of-all-political-problems-says-german-interior-minister-horst-seehofer/a-45378092>, retrieved 9 February 2022] after the protests in Chemnitz, which sees migration at the heart of society's disillusionment, shows how charged the debate about migration is in Germany. 



# Background

...


# Methods

## Hypothesis

**Are different factors relevant to the perceived threat of migrants by people with a high or low left/right orientation?**

The following factors have been identified as relevant in the literature: 



# Results

## mixed stuff (please remove this for your report!)

### overview of data

First glimpse / get summary of the loaded data: 
```{r}
glimpse(dat)
summary(dat)
psych::describe(dat)
```

```{r}
# Immigration enriching economy, culture, place to live
sel_var <- c("imbgeco", "imueclt", "imwbcnt")
getDescriptives(dataset = dat, variables = sel_var, nameAPAtable = "immigrationEnrich")
```


### identify people with extreme left / right political orientation

Example: dichotomize the `lrscale` variable (left, right scale) on the extremes:

```{r}
dat$lrscale_scaled <- scale(x = dat$lrscale, center = TRUE, scale = TRUE)
mean_upper <- mean(dat$lrscale_scaled[dat$lrscale_scaled > 1], na.rm = TRUE)
mean_lower <- mean(dat$lrscale_scaled[dat$lrscale_scaled < -1], na.rm = TRUE)
hist(dat$lrscale_scaled)
# add vertical line
abline(v=mean_upper, col = "red")
abline(v=mean_lower, col = "red")
```


```{r}
dat$lrscale_dummy <- ifelse(test = dat$lrscale_scaled > mean_upper, yes = 2, no =
         ifelse(test = dat$lrscale_scaled < mean_lower, yes = 1, no = 0))
table(dat$lrscale_dummy)
```


```{r}
tmp_desc <- dat %>%
  group_by(lrscale_dummy) %>%
  summarise(N = n(),
            meanEnrich = mean(mean_imEnrich, na.rm = TRUE),
            sdEnrich = sd(mean_imEnrich, na.rm = TRUE),
            meanAllow = mean(mean_imAllow, na.rm = TRUE),
            sdAllow = sd(mean_imAllow, na.rm = TRUE),
            meanStf = mean(mean_stf, na.rm = TRUE),
            sdStf = sd(mean_stf, na.rm = TRUE))

tmp_desc
```

Using the **stargazer** package you can create nice tables. You sometimes need to limit the number of digits using the function round(): 

```{r, results='asis'}
stargazer(round(x = tmp_desc, digits = 4), type = "html", summary = FALSE)
stargazer(x = as.data.frame(dat[, str_subset(string = colnames(dat), pattern = "mean_")]), type = "html", summary = TRUE, digits = 2)
```

### statistical tests


Combining statistical tests like **correlation** with the respective plot: 
```{r}
cor(dat$mean_stf, dat$mean_imEnrich, use = "complete")
plot(dat$mean_stf ~ dat$mean_imEnrich)
lm <- lm(formula = mean_stf ~ mean_imEnrich, data = dat)
abline(lm, col = "red")
```

Combining statistical tests like **t.test** with the respective plot: 

```{r}
dat$lrscale_dummy[dat$lrscale_dummy == 0] <- NA
boxplot(dat$mean_stf ~ dat$lrscale_dummy)
t.test(dat$mean_stf ~ dat$lrscale_dummy)
```

Based on the dichotomized `lrscale` variable compute two **multiple linear regressions**:

```{r}
# get subsets of data
dat_upper <-dat %>%
  filter(dat$lrscale_scaled > mean_upper)
dat_lower <-dat %>%
  filter(dat$lrscale_scaled < mean_lower)
# compute linear models
lm_upper <- lm(formula = mean_imEnrich ~ mean_stf + happy + agea + gndr, data = dat_upper)
lm_lower <- lm(formula = mean_imEnrich ~ mean_stf + happy + agea + gndr, data = dat_lower)
summary(lm_upper)
summary(lm_lower)
```

Again we can use the **stargazer** package, which identifies `lm_upper` for example as an lm-object in R and creates a wonderfull table for the model results: 

```{r, results='asis'}
stargazer(lm_upper, type = "html")
stargazer(lm_lower, type = "html")
```


## Quoting (please remove this for your report!)
<!-- https://blogdown-demo.rbind.io/2017/08/28/adding-citations-to-posts/ -->

How to cite literature within an rmarkdown document: 


1. Blah blah [see @rutkowski2013handbook, pp. 33-35; also @leeuw2012international, ch. 1].
2. Blah blah [@rutkowski2013handbook, pp. 33-35].
3. Blah blah [@rutkowski2013handbook; @leeuw2012international].
4. Rutkowski et al. says blah [-@rutkowski2013handbook].
5. @rutkowski2013handbook says blah.





## data analysis additional







# Discussion



# Appendix

## Test for unidimensionality

*Here we could, for example, check the unidimensionality of the summarized scales (precondition of Cronbach's Alpha).*


# References
