---
title: "my title"
author: "Julius Fenn"
date: "2021-0X-XX"
output:
  html_document:
    toc: yes
    toc_depth: 3
    number_sections: yes
  pdf_document:
    toc: yes
    toc_depth: '3'
  word_document:
    toc: yes
    toc_depth: '3'
bibliography: LibraryAll.bib
biblio-style: apalike
link-citations: yes
---


```{css, echo=FALSE}
.CSSoutput {
  background-color: white;
  border: 1px solid black;
}
```
```{r setup chunks, include=FALSE}
knitr::opts_chunk$set(class.source = "CSSoutput",
                      eval = TRUE, echo = TRUE, include = TRUE,      
                      fig.align='center', fig.show='asis',
                      size='footnotesize')

```


```{r setup, include=FALSE}
# reset R environment
rm(list=ls(all=TRUE))
graphics.off()

################
# install and load packages
################
#  if packages are not already installed, the function will install and activate them
usePackage <- function(p) {
  if (!is.element(p, installed.packages()[,1]))
    install.packages(p, dep = TRUE, repos = "http://cran.us.r-project.org")
  require(p, character.only = TRUE)
}

usePackage("haven") # load SPSS, ... data 
usePackage("tidyverse") # data cleaning and summarizing
usePackage("psych") # psychometric analysis (EFA)

## psychometric analysis
usePackage("moments") # skewness, kurtosis

## outputs
usePackage("stargazer") # create tables
rm(usePackage)
```

```{r loaddata, include=FALSE}
## change working directory
setwd("data")
## load data
dat <- haven::read_spss(file = "ESS1-9e01_1.sav")


## data preparation
dat$cntry <- NULL

# reverse coded items
dat$imsmetn <- 5 - dat$imsmetn 
dat$imdfetn <- 5 - dat$imdfetn
dat$impcntr <- 5 - dat$impcntr


## create mean scores
# > Immigration enriching economy, culture, place to live
sel_var <- c("imbgeco", "imueclt", "imwbcnt")
dat$mean_imEnrich <- rowMeans(x = dat[, sel_var], na.rm = TRUE)
# >
sel_var <- c("imsmetn", "imdfetn", "impcntr")
dat$mean_imAllow <- rowMeans(x = dat[, sel_var], na.rm = TRUE)
# >
sel_var <- str_detect(string = colnames(dat), pattern = "^trst")
dat$mean_trst <- rowMeans(x = dat[, sel_var], na.rm = TRUE)
# >
sel_var <- str_detect(string = colnames(dat), pattern = "^stf")
dat$mean_stf <- rowMeans(x = dat[, sel_var], na.rm = TRUE)
dat$sd_stf <- dat %>%
  select(matches(match = "^stf")) %>%
  apply(.,1, sd, na.rm = TRUE)
setwd("..")
```











# Introduction

Migration 'mother of all political problems,' says German Interior Minister Horst Seehofer. Hort Seehofer's 2018 statement^[see <https://www.dw.com/en/migration-mother-of-all-political-problems-says-german-interior-minister-horst-seehofer/a-45378092>, retrieved 9 February 2022] after the protests in Chemnitz, which sees migration at the heart of society's disillusionment, shows how charged the debate about migration is in Germany. 



# Background

...


# Methods

## Hypothesis

**Are different factors relevant to the perceived threat of migrants by people with a high or low left/right orientation?**

The following factors have been identified as relevant in the literature: 



# Results

## mixed stuff (please remove this for your report!)

First glimpse at the loaded data: 
```{r}
glimpse(dat)
summary(dat)
```


Combining statistical tests like cor.test or t.test with the respective plot: 
```{r}
cor(dat$mean_stf, dat$mean_imEnrich, use = "complete")
plot(dat$mean_stf, dat$mean_imEnrich)
```

Check the assumptions of every test procedure you are using!
```{r}
t.test(dat$mean_stf ~ dat$gndr)
boxplot(dat$mean_stf ~ dat$gndr)
```


## Quoting (please remove this for your report!)
<!-- https://blogdown-demo.rbind.io/2017/08/28/adding-citations-to-posts/ -->

How to cite literature within an rmarkdown document: 


1. Blah blah [see @rutkowski2013handbook, pp. 33-35; also @leeuw2012international, ch. 1].
2. Blah blah [@rutkowski2013handbook, pp. 33-35].
3. Blah blah [@rutkowski2013handbook; @leeuw2012international].
4. Rutkowski et al. says blah [-@rutkowski2013handbook].
5. @rutkowski2013handbook says blah.

## data preperation 

Aim to dichotomize the `lrscale` variable:

```{r}
dat$lrscale_scaled <- scale(x = dat$lrscale, center = TRUE, scale = TRUE)
mean_upper <- mean(dat$lrscale_scaled[dat$lrscale_scaled > 1], na.rm = TRUE)
mean_lower <- mean(dat$lrscale_scaled[dat$lrscale_scaled < -1], na.rm = TRUE)
hist(dat$lrscale_scaled)
# add vertical line
abline(v=mean_upper, col = "red")
abline(v=mean_lower, col = "red")
# save subsets
dat_upper <-dat %>%
  filter(dat$lrscale_scaled > mean_upper)
dat_lower <-dat %>%
  filter(dat$lrscale_scaled < mean_lower)
```


## data analysis descriptive

```{r}
dat$lrscale_dummy <- ifelse(test = dat$lrscale_scaled > mean_upper, yes = 2, no =
         ifelse(test = dat$lrscale_scaled < mean_lower, yes = 1, no = 0))
table(dat$lrscale_dummy)
```


```{r}
tmp_desc <- dat %>%
  group_by(lrscale_dummy) %>%
  summarise(N = n(),
            meanEnrich = mean(mean_imEnrich, na.rm = TRUE),
            sdEnrich = sd(mean_imEnrich, na.rm = TRUE),
            meanAllow = mean(mean_imAllow, na.rm = TRUE),
            sdAllow = sd(mean_imAllow, na.rm = TRUE),
            meanStf = mean(mean_stf, na.rm = TRUE),
            sdStf = sd(mean_stf, na.rm = TRUE),
            meanTrst = mean(mean_trst, na.rm = TRUE),
            sdTrst = sd(mean_trst, na.rm = TRUE))

tmp_desc
```

Using the **stargazer** package you can create nice tables. You sometimes need to limit the number of digits using the function round(): 

```{r, results='asis'}
stargazer(round(x = tmp_desc, digits = 4), type = "html", summary = FALSE)
stargazer(x = as.data.frame(dat[, str_subset(string = colnames(dat), pattern = "mean_")]), type = "html", summary = TRUE, digits = 2)

knitr::kable(mtcars[1:2, 1:2], "html")

```




## data analysis main

```{r}
lm_upper <- lm(formula = mean_imEnrich ~ mean_trst + mean_stf + happy + agea + gndr, data = dat_upper)
lm_lower <- lm(formula = mean_imEnrich ~ mean_trst + mean_stf + happy + agea + gndr, data = dat_lower)
summary(lm_upper)
summary(lm_lower)
```

Again we can use the **stargazer** package, which identifies `lm_upper` for example as an lm-object in R and creates a wonderfull table for the model results: 

```{r, results='asis'}
stargazer(lm_upper, type = "html")
stargazer(lm_upper, type = "html")
```

```{r}
ggplot(mpg, aes(displ, hwy, colour = class)) + 
  geom_point()
```



## data analysis additional







# Discussion



# Appendix

## Test for unidimensionality

Here we could apply a self-wirtten function for example to check the reliability and amount of explained variance for the first factor: 

```{r}
corr_rel_EFA <- function(constructlist = NULL, constnum = NULL,
                         data = NULL,
                         nfacs = 1){
  ### correlation measures:
  # spearman
  cor_mat <- cor(data[,constructlist[[constnum]]],
                 use = "pairwise.complete.obs",
                 method = "spearman")

  ### reliability measures:
  # Cronbachs
  rel_cronbach <- psych::alpha(cor_mat)

  ### EFA (PAF):
  fit_efa <- fa(r = data[,constructlist[[constnum]]], nfactors = nfacs,
                rotate = "Promax", fm = "pa", max.iter = 500)


  ### return objects as list
  return_list <- list(round(x = cor_mat, digits = 2),
                      rel_cronbach,
                      fit_efa
  )
  names(return_list) <- c("Cor: Spearman",
                          "Reliability: Cronbach",
                          "fit EFA (PAF)")

  # > print
  cat("mean inter-item-correlation (Spearman):",
      round(x = mean(colMeans(x = cor_mat)), digits = 2), "\n\n")

  cat("Cronbachs Alpha:",
      round(x = rel_cronbach$total[[1]], digits = 2), "\n\n")

  cat("EFA (PAF) variance accounted first factor:",
      round(x = fit_efa$Vaccounted[2], digits = 2), "for", nfacs, "factors", "\n")
  tmpKMO <- psych::KMO(cor_mat)
  if(any(tmpKMO$MSAi < .6)){
    cat("KMO criteria is to low (< .6) for:", "\n",
        names(tmpKMO$MSAi[tmpKMO$MSAi < .6]), "\n",
        "mean KMO:", round(x = tmpKMO$MSA, digits = 2), "\n")
  }
  #

  return(return_list)
}


constructs_list <- list()
(vars_tmp <-  c("imbgeco", "imueclt", "imwbcnt"))
constructs_list[[1]] <- vars_tmp
(vars_tmp <-  str_subset(string = colnames(dat), pattern = "^trst"))
constructs_list[[2]] <- vars_tmp

tmp <- corr_rel_EFA(constructlist = constructs_list, constnum = 1, data = dat, nfacs = 1)
tmp
tmp <- corr_rel_EFA(constructlist = constructs_list, constnum = 2, data = dat, nfacs = 2)
tmp
```


That we have a possible 2 factor solution (but relative low eigenvalue for the second factor!) for the trust variables become clear if you look at the correlation plots: 
```{r}
psych::corPlot(dat[, constructs_list[[2]]],numbers=TRUE,stars=TRUE)
psych::corPlot(dat[, c("trstprl", "trstplt", "trstprt", "trstep",  "trstun")],numbers=TRUE,stars=TRUE)
efa_parallel <- fa.parallel(x = dat[, constructs_list[[2]]], fa = "fa",n.iter=50)
efa_parallel
```

A parallel analysis even suggesting a 2-3 factor solution: 
```{r}
efa_parallel <- fa.parallel(x = dat[, constructs_list[[2]]], fa = "fa",n.iter=50)
efa_parallel
```


# References
